Modeling Intervention Strategies for United States TB Control

### 2013-08-25 in_progress/presentations/paper/draft.tex
working on paper

\documentclass{amsart}
\usepackage[pdftex]{graphicx}

@@

Further implementations of the \texttt{C++} model were made that deviated from
the basic Hill Model. Specifically, the acute latent compartment in the basic
Hill Model is a vestigal necessity of compartmental modeling and not reflective
of the biology of tuberculosis, and an agent based model was implemented that
more closely respects the biology of tuberculosis. Results from this model or
the strict Hill model did not differ significantly. 
\section{Results}

\subsection{Basic Population Breakdown}
The additional tracking capabilities offered several key insights into US TB
dynamics. Figure INSERT FIGURE HERE shows the yearly incidence of US TB, broken
down into infection source. It can be seen that the majority of the US TB load
is driven by activations of FB LTBI, followed by USB LTBI activations. This data
further agrees with the conclusions drawn by Hill, Becerra, and Castro about the
necessity of LTBI treatment in any valuable intervention strategy. Further,
figure INSERT FIGURE HERE shows a similarly sourced plot, but analyzing the
final US HCS costs due to TB. One can see that in this plot, roughly half of the
US TB HCS costs are due to activations of LTBI. Note that both of these plots
underestimate the impact LTBI activations play in the spread of TB, as every
LTBI activation to infectious TB contributes not only to incidence and costs
directly, but also indirectly by causing additional future cases, which is not
captured in these graphs. 
\subsubsection{Basic Reproduction Number}
The basic reproduction number of FB or USB cases of infectious TB was also
estimated by this system. From a theoretical perspective, we can think of the
total number of secondary infections over 100 years due to a FB or USB
infectious TB case as describing a geometric series in a large population.
Presuming there are no overlaps in infectious contacts, if a single case of
infectious TB in either population infects $p_f, p_u$ new cases in one year,
respectively, then we can say that over the course of 100 years, the total
number of cases infected will follow a geometric series. This analysis predicts
that over 100 years, one USB infection will lead to $1.03$ subsequent infections,
whereas one FB infection will lead to $.64$ subsequent infections. The full calculations
are included in the appendix.  Experimentally, these data were also analyzed, 
with results of $1.04$ and $.8$, respectively.

\subsection{Intervention Analysis}
The primary interventions analyzed by the extended Hill were those that analyzed
curing various percentages of entering LTBI cases. Four indicative percentages
chosen were $20\%$, $35\%$, $50\%$, and $65\%$. Note that the Hill model does
not distinguish documented immigration from undocumented immigration, and as
such estimates of entering LTBI cure rates higher than $65\%$ become much more
difficult to achieve. It was seen that in this model as well as in the basic
Hill, no analyzed intervention predicted elimination by 2100. In order to obtain
elimination by 2100, at least 95\% of entering LTBI cases had to be cured, which
is practically impossible. However, it was seen that curing entering cases of
LTBI resulted in a net US HCS cost per case averted of \$67,722.96 at 2025, \$28,759.00 at 2050,
and \$17,967.39 at 2100 (assuming 20\% reduction with one cure costing \$XXXX; see Table
\ref{tab:cpcaRes} for other percentages). Further, it was also found that the relationship
between total incidence at 2100 and percentage of incoming LTBI cases cured was linear, and
from this estimates were made of the yearly average US HCS savings garnered by
curing one case of entering LTBI over the time scale 2000 to 2025, 2000 to 2050,
and 2000 to 2100. This value peaked at \$1.026 million at 2100 (20\% reduction). This illustrates that it
would be cost saving to cure cases of LTBI at the cost of \$1.026 million ``2000'' dollars over
the time period 2000-2100. These intervention strategies also resulted in 23,870; 41,952;
and 60,189 fewer cases of TB seen in the US, and 2,056; 3,614; and 5,185 fewer TB
deaths, for 20\%, 35\%, and 50\% reduction, respectively.

@@

\subsection{Sensitivity Analysis}
\section{Discussion}
These results confirm the hypothesis that curing incoming LTBI rates is a
necessary step towards elimination and indicate that it is a cost effective
option. 
\end{document}

>>>>

\documentclass{amsart}
\usepackage[pdftex]{graphicx}
\graphicspath{{figures/}}

@@

Further implementations of the \texttt{C++} model were made that deviated from
the basic Hill Model. Specifically, the acute latent compartment in the basic
Hill Model is a vestigal necessity of compartmental modeling and not reflective
of the biology of tuberculosis, and an agent based model was implemented that
more closely respects the biology of tuberculosis. Results from this model or
the strict Hill model did not differ significantly. 
\section{Results}
\subsection{Basic Population Breakdown}
The additional tracking capabilities offered several key insights into US TB
dynamics. Figure~\ref{fig:incPlotSourced} shows the yearly incidence of US TB, broken
down into infection source. It can be seen that the majority of the US TB load
is driven by activations of FB LTBI, followed by USB LTBI activations. This data
further agrees with the conclusions drawn by Hill, Becerra, and Castro about the
necessity of LTBI treatment in any valuable intervention strategy. 
\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=.5]{incPlotSourced}
  \end{center}
  \caption{Sourced yearly incidence data generated by the extended Hill Model.}
  \label{fig:incPlotSourced}
\end{figure}
Further, figure~\ref{fig:costPlotSourced} shows a similarly sourced plot, but analyzing the
final US HCS costs due to TB. One can see that in this plot, roughly half of the
US TB HCS costs are due to activations of LTBI. In this data, it is appropriate
to think of the costs due to acute LTBI are due to exogenous re-inection,
whereas active TB costs due to chronic LTBI are activations of LTBI into active
TB. 
\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=.35]{costPlotSourced}
  \end{center}
  \caption{Sourced US HCS economic TB load. Note that this data only illustrates
    the load due to treating active TB, but illustrates where the infections
    driving this cost come from.}
  \label{fig:costPlotSourced}
\end{figure}
Note that both of these plots underestimate the impact LTBI activations play in
the spread of TB, as every LTBI activation to infectious TB contributes not only
to incidence and costs directly, but also indirectly by causing additional
future cases, which is not captured in these graphs. Further, there is also US
HCS costs due to treating LTBI, which is not illustrated in these graphs. 
\subsubsection{Basic Reproduction Number}
The basic reproduction number of FB or USB cases of infectious TB was also
estimated by this system. From a theoretical perspective, we can think of the
total number of secondary infections over 100 years due to a FB or USB
infectious TB case as describing a geometric series in a large population.
Presuming there are no overlaps in infectious contacts, if a single case of
infectious TB in either population infects $p_f, p_u$ new cases in one year,
respectively, then we can say that over the course of 100 years, the total
number of cases infected will follow a geometric series. This analysis predicts
that over 100 years, one USB infection will lead to $1.03$ subsequent infections,
whereas one FB infection will lead to $.64$ subsequent infections. The full calculations
are included in the appendix.  Experimentally, these data were also analyzed, 
with results of $1.04$ and $.8$, respectively. WHAT CAUSES THIS DIFFERENCE? WE
NEED TO ANSWER THAT.

\subsection{Intervention Analysis}
The primary interventions analyzed by the extended Hill were those that analyzed
curing various percentages of entering LTBI cases. Four indicative percentages
chosen were $20\%$, $35\%$, $50\%$, and $65\%$. Note that the Hill model does
not distinguish documented immigration from undocumented immigration, and as
such estimates of entering LTBI cure rates higher than $65\%$ become much more
difficult to achieve. It was seen that in this model as well as in the basic
Hill, no analyzed intervention predicted elimination by 2100. In order to obtain
elimination by 2100, at least 95\% of entering LTBI cases had to be cured, which
is practically impossible. However, it was seen that curing entering cases of
LTBI resulted in a net US HCS cost per case averted of \$67,722.96 at 2025, \$28,759.00 at 2050,
and \$17,967.39 at 2100 (assuming 20\% reduction with one cure costing \$XXXX; see Table
\ref{tab:cpcaRes} for other percentages). Given the variable nature of LTBI
treatment cost, the model code is extendible such that a user can adjust these
costs themselves to explore more specific methods of curing entering LTBI.
Further, it was also found that the relationship between total incidence at 2100
and percentage of incoming LTBI cases cured was linear, and from this estimates
were made of the yearly average US HCS savings garnered by curing one case of
entering LTBI over the time scale 2000 to 2025, 2000 to 2050, and 2000 to 2100.
This value peaked at \$1.026 million at 2100 (20\% reduction). This illustrates
that it would be cost saving to cure cases of LTBI at the cost of \$1.026
million ``2000'' dollars over the time period 2000-2100. These intervention
strategies also resulted in 23,870; 41,952; and 60,189 fewer cases of TB seen in
the US, and 2,056; 3,614; and 5,185 fewer TB deaths, for 20\%, 35\%, and 50\%
reduction, respectively.

@@

\subsection{Sensitivity Analysis}
\section{Discussion}
These results confirm the hypothesis that curing incoming LTBI rates is a
necessary step towards elimination and indicate that it is a cost effective
option. 
\subsection{Future Work}
This work could be extended by examining different classes of interventions or
more accurately estimating intervention cost with the deterministic extended
Hill. Further work could also be done with the agent-based Hill Model, by using
it to examine the effect contact structure plays on US TB incidence levels or to
examine the effects drug-resistant TB will have on US TB dynamics. 
\end{document}
